<!-- question-type: inclass -->
### Exercise 5: Comparing Rule-of-Thumb Segmentation to K-Means

In Exercise 2, you created four employee groups using simple rules based on `job_satisfaction` and `distance_from_home`.

Now we will use **K-means clustering** on those same two variables and compare the result.

**(a)** Create a new dataset called `df_scaled_small` that keeps only the scaled versions of:

* `job_satisfaction`
* `distance_from_home`

from `df_scaled`.

```{r}
#| eval: false
df_scaled_small <-
  df_scaled |>
  select(YOUR_CODE, YOUR_CODE)
```

**(b)** Run K-means clustering on `df_scaled_small` using **4 clusters**.

```{r}
#| eval: false
set.seed(123)

kmeans_small <-
  kmeans(
    YOUR_CODE,
    centers = YOUR_CODE,
    nstart = 10
  )
```

**(c)** Use the code below to visualise the clustering result.

```{r}
#| eval: false
fviz_cluster(
  kmeans_small,
  data = df_scaled_small,
  ggtheme = theme_minimal(),
  main = "2-variable K-means clustering"
)
```

**(d)** What do you notice about the shape of the clusters? Do they look very different from the rule-of-thumb groups in Exercise 2?

**(e)** The code below combines the original employee data with the K-means cluster assignments.

Using `df_compare`, calculate the number of employees and the attrition rate in each cluster.

```{r}
#| eval: false
df_compare <-
  df |>
  select(attrition, job_satisfaction, distance_from_home) |>
  mutate(cluster = as.factor(kmeans_small$cluster))

df_compare |>
  YOUR_CODE_HERE
```

**(f)** Compared with the rule-of-thumb segmentation, did K-means appear to add much in this two-variable setting? Explain briefly.

<!-- BEGIN PROFILE:r-teaching-guide -->

::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

* Run K-means clustering on a simple two-variable dataset.
* Compare a data-driven clustering result to a hand-built rule-of-thumb segmentation.
* Summarise cluster sizes and attrition rates using grouped summaries.
* Reflect on whether K-means adds meaningful insight in a simple low-dimensional setting.

✅   **Core Concepts to Highlight**

* K-means can be run on the same two variables used in the earlier rule-based segmentation.
* Setting `centers = 4` makes the comparison to the four rule-based groups especially direct.
* In a simple two-variable setting, K-means may produce clusters that look quite similar to hand-built buckets.
* The value of K-means is not automatic; whether it adds insight depends on the data and the problem.
* Cluster summaries should be interpreted using proportions such as attrition rates, not counts alone.

💬 **Suggested In-Class Prompts** (if needed)

* “Why might we deliberately use the same two variables here as in Exercise 2?”
* “Does the clustering plot suggest something meaningfully different from the simple four-group rule?”
* “If K-means looks similar to the rule-based segmentation here, is that a problem?”
* “What do the attrition rates tell us that the plot alone does not?”

📌 **Common Misunderstandings**

* Assuming K-means must always outperform a simple hand-built segmentation
* Treating the visual cluster plot as sufficient without checking cluster summaries
* Focusing only on cluster size and forgetting to compare attrition rates
* Thinking similarity between the rule-based groups and K-means means the algorithm failed, when it may simply indicate that the simple structure was already capturing the main pattern

:::

:::

:::

<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->

::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a) Create a new dataset called `df_scaled_small` that keeps only the scaled versions of `job_satisfaction` and `distance_from_home` from `df_scaled`.**

```{r}
#| eval: true
df_scaled_small <-
  df_scaled |>
  select(job_satisfaction, distance_from_home)
```

**(b) Run K-means clustering on `df_scaled_small` using 4 clusters.**

```{r}
#| eval: true
set.seed(123)

kmeans_small <-
  kmeans(
    df_scaled_small,
    centers = 4,
    nstart = 10
  )
```

**(c) Use the code below to visualise the clustering result.**

```{r}
#| eval: true
fviz_cluster(
  kmeans_small,
  data = df_scaled_small,
  geom = "point",
  ggtheme = theme_minimal(),
  main = "2-variable K-means clustering"
)
```

**(d) What do you notice about the shape of the clusters? Do they look very different from the rule-of-thumb groups in Exercise 2?**

A strong answer should note that the clusters may look fairly similar to the earlier four-group rule-based segmentation. In a simple two-variable setting, K-means may not add dramatically different structure.

**(e) Using `df_compare`, calculate the number of employees and the attrition rate in each cluster.**

```{r}
#| eval: true
df_compare <-
  df |>
  select(attrition, job_satisfaction, distance_from_home) |>
  mutate(cluster = as.factor(kmeans_small$cluster))

df_compare |>
  group_by(cluster) |>
  summarise(
    n_employees    = n(),
    attrition_rate = mean(attrition == "Yes"),
    job_sat        = mean(job_satisfaction),
    distance       = mean(distance_from_home)
  )
```

**(f) Compared with the rule-of-thumb segmentation, did K-means appear to add much in this two-variable setting? Explain briefly.**

The four clusters line up closely with the four combinations of lower versus higher job satisfaction and shorter versus longer commute distance. For example, clusters 1 and 2 both have lower average job satisfaction, while clusters 3 and 4 both have higher average job satisfaction. Clusters 2 and 3 both have longer average commute distances, while clusters 1 and 4 have shorter average commute distances.

The attrition rates also follow a familiar pattern. The highest attrition appears in the cluster with lower satisfaction and longer commute (cluster 2), while the lowest attrition appears in the cluster with higher satisfaction and shorter commute (cluster 4). The other two clusters sit in between. This suggests that in a two-variable setting, K-means may not add dramatically different insight beyond what a transparent rule-of-thumb segmentation already captured.

That does not mean K-means is useless. Rather, it shows that when the structure is simple and the variables are few, a hand-built segmentation can already do a reasonable job of describing the main pattern in the data.

:::

:::

:::

<!-- END PROFILE:r-solutions -->
