<!-- question-type: inclass -->
### Exercise 6: Building a Richer K-Means Model

Your analytics team now wants to build a richer clustering model using all five scaled variables in `df_scaled`:

* `job_satisfaction`
* `monthly_income`
* `years_at_company`
* `distance_from_home`
* `age`

One challenge remains: we still need to choose a value for `k`, the number of clusters.

A colleague suggests using the **elbow method** as a starting point.

**(a)** Explain what the elbow method is trying to do.

**(b)** Run the code below to create an elbow plot for K-means clustering on `df_scaled`.

```{r}
#| eval: false
fviz_nbclust(
  df_scaled,
  kmeans,
  method = "wss",
  k.max = 10,
  nstart = 10
)
```

**(c)** Based on the elbow plot, what value of `k` would you choose? Briefly explain your answer.

**(d)** Run K-means clustering on `df_scaled` using your chosen value of `k`.

```{r}
#| eval: false
set.seed(123)

kmeans_full <-
  kmeans(
    YOUR_CODE,
    centers = YOUR_CODE,
    nstart = 10
  )
```

**(e)** Use the code below to visualise the clustering result.

```{r}
#| eval: false
fviz_cluster(
  kmeans_full,
  data = df_scaled,
  ggtheme = theme_minimal(),
  main = "5-variable K-means clustering"
)
```

**(f)** What can the plot in (e) tell you about the clustering result? What can it not tell you on its own?

**(g)** The code below combines the original employee data with the cluster assignments from your 5-variable K-means model.

Using `df_compare_full`, calculate:

* the number of employees in each cluster
* the attrition rate in each cluster
* the average value of each clustering variable in each cluster

```{r}
#| eval: false
df_compare_full <-
  df |>
  select(
    attrition,
    job_satisfaction,
    monthly_income,
    years_at_company,
    distance_from_home,
    age
  ) |>
  mutate(cluster = as.factor(kmeans_full$cluster))

df_compare_full |>
  YOUR_CODE_HERE
```

<!-- BEGIN PROFILE:r-teaching-guide -->

::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

* Use the elbow method to support a choice of `k` in a richer K-means model.
* Run and inspect a 5-variable clustering model.
* Recognize that choosing `k` involves judgment rather than a fully mechanical rule.
* Summarise richer clusters using attrition rates and cluster means.

✅   **Core Concepts to Highlight**

* Moving from two variables to five gives K-means more scope to uncover broader employee profiles.
* The elbow method looks at how within-cluster variation changes as `k` increases.
* The goal is not to find a magically perfect answer, but to identify a defensible value of `k`.
* Cluster plots are helpful, but in higher dimensions they are only partial summaries.
* Cluster summaries are needed to understand what the clusters actually represent.

💬 **Suggested In-Class Prompts** (if needed)

* “Why might the richer five-variable model reveal patterns the two-variable model could not?”
* “What is the elbow method trying to balance?”
* “Does the elbow plot give one objectively correct answer, or does it still require judgment?”
* “What can the cluster plot show us, and what do we still need the summary table for?”

📌 **Common Misunderstandings**

* Thinking the elbow plot always reveals a perfectly obvious value of `k`
* Treating the elbow method as a strict rule rather than a decision aid
* Assuming the cluster plot is enough to understand five-dimensional clusters
* Forgetting to compare attrition rates and average characteristics together when interpreting clusters

:::

:::

:::

<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->

::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a) A colleague suggests using the elbow method. Explain what the elbow method is trying to do.**

The elbow method helps analysts choose a reasonable number of clusters by plotting how within-cluster variation changes as k increases. As more clusters are added, this variation will always fall, but after a certain point the improvement becomes much smaller. The analyst then looks for the "elbow" in the plot, where adding another cluster no longer gives a large reduction in within-cluster variation.

**(b) Run the code below to create an elbow plot for K-means clustering on `df_scaled`.**

```{r}
#| eval: true
fviz_nbclust(
  df_scaled,
  kmeans,
  method = "wss",
  k.max = 10,
  nstart = 10
)
```

**(c) Based on the elbow plot, what value of `k` would you choose? Briefly explain your answer.**

A strong answer should choose a defensible value of `k` and justify it using the shape of the elbow plot. The key idea is that improvements become smaller after a certain point, even if the elbow is not perfectly sharp. We went with `k = 4`

**(d) Run K-means clustering on `df_scaled` using your chosen value of `k`.**

```{r}
#| eval: true
set.seed(123)

kmeans_full <-
  kmeans(
    df_scaled,
    centers = 4,
    nstart = 10
  )
```

**(e) Use the code below to visualise the clustering result.**

Students should produce the cluster plot and use it as one input into interpretation, alongside the cluster summaries.

```{r}
#| eval: true
fviz_cluster(
  kmeans_full,
  data = df_scaled,
  ggtheme = theme_minimal(),
  main = "5-variable K-means clustering"
)
```

**(f) What can this plot tell you about the clustering result? What can it not tell you on its own?**

The plot can give a broad sense of whether clusters appear reasonably separated, but it cannot fully describe the cluster profiles in a five-variable setting. For that, we still need summary statistics for each cluster.

**(g) Using `df_compare_full`, calculate the number of employees in each cluster, the attrition rate in each cluster, and the average value of each clustering variable in each cluster.**

Follow the same logic as in Exercise 5, using `group_by(cluster)` and `summarise()` to calculate cluster sizes, attrition rates, and means for the five variables.


```{r}
#| eval: true
df_compare_full <-
  df |>
  select(
    attrition,
    job_satisfaction,
    monthly_income,
    years_at_company,
    distance_from_home,
    age
  ) |>
  mutate(cluster = as.factor(kmeans_full$cluster))

df_compare_full |>
  group_by(cluster) |>
  summarise(
    n_employees    = n(),
    attrition_rate = mean(attrition == "Yes"),
    job_sat        = mean(job_satisfaction),
    income         = mean(monthly_income),
    years          = mean(years_at_company),
    distance       = mean(distance_from_home),
    age            = mean(age)
  )
```

:::

:::

:::

<!-- END PROFILE:r-solutions -->
