<!-- question-type: prepare -->
### Exercise 2: A Simple Rule-of-Thumb Employee Segmentation

Suppose your analytics team wants to create a quick first-pass classification of employees using two variables:

* `job_satisfaction`
* `distance_from_home`

For this exercise, define:

* **lower satisfaction** as a score **below 2**
* **longer commute** as living **10 km or more** from work

This gives four possible employee groups:

* Higher satisfaction, shorter commute
* Higher satisfaction, longer commute
* Lower satisfaction, shorter commute
* Lower satisfaction, longer commute

**(a)** Why might an analytics team begin with a simple rule-of-thumb segmentation before trying a more data-driven method?

**(b)** Complete the code below to classify each employee into one of the four segments.

```{r}
#| eval: false
df_rule <- 
  df |>
  mutate(
    rule_segment = case_when(
      YOUR_CONDITION ~ "Higher satisfaction, shorter commute",
      YOUR_CONDITION ~ "Higher satisfaction, longer commute",
      YOUR_CONDITION ~ "Lower satisfaction, shorter commute",
      .default = "Lower satisfaction, longer commute"
    )
  )
```

**(c)** For each segment, calculate:

* the number of employees
* the attrition rate

```{r}
# Write your answer here
```

**(d)** Which segment or segments appear to have the highest attrition?

**(e)** What are the advantages of this rule-of-thumb approach for an analytics team? What kinds of patterns might it fail to capture?

<!-- BEGIN PROFILE:r-teaching-guide -->

::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

* Build a simple rule-based employee segmentation using fixed thresholds.
* Interpret differences in attrition across segments.
* Recognize the strengths and limitations of hand-built segmentation rules.

✅   **Core Concepts to Highlight**

* Segmentation as a way of grouping observations into meaningful categories
* Rule-of-thumb segmentation as simple, transparent, and easy to communicate
* Fixed thresholds as analyst choices that shape the resulting groups
* Attrition rate as a useful summary for comparing segments
* The idea that simple rules can be useful, but may miss more complex patterns in the data

💬 **Suggested In-Class Prompts** (if needed)

* “Why might a team start with a simple segmentation rule before trying something more data-driven?”
* “What do we gain by using thresholds that are easy to explain?”
* “What might this rule-based approach miss about employee differences?”
* “If two segments have similar attrition rates, does that mean they are really the same kind of employee?”

📌 **Common Misunderstandings**

* Thinking the chosen cutoffs are objectively correct rather than simple analyst-made rules
* Assuming a clear rule-based segmentation must capture all important patterns
* Treating groups with similar attrition as identical, even if they differ on underlying characteristics
* Forgetting to use proportions rather than only counts when comparing attrition across groups

:::

:::

:::

<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->

::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a) Why might an analytics team begin with a simple rule-of-thumb segmentation before trying a more data-driven method?**

* **Easy to understand**: Simple rules are transparent and easy to explain to others.
* **Fast to implement**: Analysts can create an initial segmentation quickly.
* **Useful benchmark**: A hand-built rule gives a baseline that can later be compared to more data-driven methods.
* **Business intuition**: The chosen variables may reflect concerns that managers already care about.

**(b) Complete the code below to classify each employee into one of the four segments.**

```{r}
#| eval: false
df_rule <- 
  df |>
  mutate(
    rule_segment = case_when(
      job_satisfaction >= 2 & distance_from_home < 10 ~ "Higher satisfaction, shorter commute",
      job_satisfaction >= 2 & distance_from_home >= 10 ~ "Higher satisfaction, longer commute",
      job_satisfaction < 2 & distance_from_home < 10 ~ "Lower satisfaction, shorter commute",
      .default = "Lower satisfaction, longer commute"
    )
  )
```

**(c) For each segment, calculate the number of employees and the attrition rate.**

```{r}
#| eval: false
df_rule |>
  group_by(rule_segment) |>
  summarise(
    n_employees = n(),
    attrition_rate = mean(attrition == "Yes")
  )
```

**(d) Which segment or segments appear to have the highest attrition?**

A strong answer should identify the segment or segments with the highest attrition rate based on the summary table, rather than relying on counts alone.

**(e) What are the advantages of this rule-of-thumb approach for an analytics team? What kinds of patterns might it fail to capture?**

* **Advantages**: The approach is simple, transparent, quick to implement, and easy to communicate.
* **What it might miss**: It depends heavily on the chosen thresholds, may ignore more subtle combinations of variables, and may fail to capture patterns that are not well described by simple cutoffs.
* **Big picture**: This is useful as a first-pass segmentation, but it may be cruder than a more data-driven clustering approach.

:::

:::

:::

<!-- END PROFILE:r-solutions -->
