<!-- question-type: inclass -->
### Exercise 4: Preparing the Data for K-Means Clustering

In the remaining exercises, we will use the following five numeric variables to build and compare K-means clustering models:

* `job_satisfaction`
* `monthly_income`
* `years_at_company`
* `distance_from_home`
* `age`

In this exercise, we prepare these five variables for the clustering analysis that follows.

**(a)** Create a dataset that contains the five variables listed above.

```{r}
#| eval: false
df_kmeans <- 
  df |>
  select(
    job_satisfaction,
    monthly_income,
    years_at_company,
    YOUR_CODE,
    YOUR_CODE
  )
```

**(b)** Complete the code below to create a recipe for `df_kmeans` and add a step that normalises all numeric predictors.

```{r}
#| eval: false
rec <- 
  YOUR_CODE(~ ., data = df_kmeans) |>
  YOUR_CODE(all_numeric_predictors())
```

**(c)** Explain what the code in part (b) is doing.

**(d)** Complete the code below to implement the preprocessing steps from the recipe in part (b) and apply them to the data.

```{r}
#| eval: false
df_scaled_5 <- 
  rec |>
  YOUR_CODE() |>
  YOUR_CODE(new_data = NULL)
```

**(e)** Explain what the two functions you used in part (d) are doing.

**(f)** Run the code below to verify that the preprocessing achieves the desired result. What do you notice about the means and standard deviations?

```{r}
#| eval: false
df_scaled |>
  summarise(
    across(
      everything(),
      list(mean = mean, sd = sd)
    )
  )
```

*Write your answer here.*

<!-- BEGIN PROFILE:r-teaching-guide -->

::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

* Prepare a numeric dataset for K-means clustering using a simple `recipes` workflow.
* Explain in plain language what `recipe()`, `step_normalize()`, `prep()`, and `bake()` are doing.
* Verify that preprocessing produced the intended result.

✅   **Core Concepts to Highlight**

* The five numeric variables prepared here will be used in the remaining clustering exercises.
* `recipe()` defines a preprocessing plan.
* `step_normalize()` adds a step that rescales numeric predictors onto a comparable scale.
* `prep()` learns what is needed to carry out the preprocessing steps.
* `bake()` applies those steps and returns the processed data.
* Good analytics practice includes checking whether preprocessing actually worked.

💬 **Suggested In-Class Prompts** (if needed)

* “Why are we setting up the five-variable dataset now rather than rebuilding it in each later exercise?”
* “What is the difference between defining a preprocessing plan and actually carrying it out?”
* “Why might it be useful to separate `prep()` from `bake()` in a data workflow?”
* “After scaling, what should we expect the means and standard deviations to look like?”

📌 **Common Misunderstandings**

* Thinking `recipe()` changes the data immediately, when it only defines the preprocessing plan
* Thinking `step_normalize()` runs by itself, when it only adds a step to the plan
* Treating `prep()` and `bake()` as interchangeable, rather than as learning and applying stages
* Forgetting that the final summary check is there to confirm the preprocessing worked as intended

:::

:::

:::

<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->

::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

**(a) The code below creates a smaller dataset containing only the five variables we will use. Complete the `select()` step.**

```{r}
#| eval: true
df_kmeans <- 
  df |>
  select(
    job_satisfaction,
    monthly_income,
    years_at_company,
    distance_from_home,
    age
  )
```

**(b) Complete the code below to create a recipe for `df_kmeans` and add a step that normalises all numeric predictors.**

```{r}
#| eval: true
rec <- 
  recipe(~ ., data = df_kmeans) |>
  step_normalize(all_numeric_predictors())
```

**(c) Explain in plain language what the code in part (b) is doing.**

This code creates a preprocessing plan for the variables in `df_kmeans` and adds a step that will normalise all numeric predictors so they are on a comparable scale.

**(d) Complete the code below to implement the preprocessing steps from the recipe in part (b) and apply them to the data.**

```{r}
#| eval: true
df_scaled <- 
  rec |>
  prep() |>
  bake(new_data = NULL)
```

**(e) Explain in plain language what the two functions you used in part (d) are doing.**

`prep()` works out what is needed to carry out the preprocessing steps in the recipe. In this case, it calculates the information needed to normalise each variable. `bake(new_data = NULL)` then applies those prepared steps to the data and returns the processed dataset.

**(f) Run the code below to verify that the preprocessing achieves the desired result. What do you notice about the means and standard deviations?**

The means are now close to 0 and the standard deviations are close to 1 for each variable. This indicates that the variables have been normalised successfully.

```{r}
#| eval: true
df_scaled |>
  summarise(
    across(
      everything(),
      list(mean = mean, sd = sd)
    )
  )
```

:::

:::

:::

<!-- END PROFILE:r-solutions -->
