## The Business Challenge

### The Topic: Why are employees leaving, and can segmentation help us understand different attrition profiles? {.unnumbered}

Every time an employee leaves, companies incur real costs through recruitment, training, disruption to teams, and lost experience. For an HR analytics team, an important first step is to understand whether **different types of employees show different patterns of attrition**.

In this tutorial, we use segmentation to explore that question. We begin with a simple **rule-of-thumb grouping** based on a small number of variables, and then compare it to **K-means clustering**, a more data-driven way of grouping employees with similar characteristics. This helps us move beyond one-off cases and identify broader employee profiles, which is a common task in **descriptive analytics**.

Our goal is not to build a formal prediction model or identify the causes of attrition. Instead, the goal is to explore whether different segmentation approaches reveal useful patterns that an analytics team could investigate further.

### The Data: HR Analytics Data {.unnumbered}

We will work with an anonymised dataset from a company's HR department. It contains information on 1,470 employees, including:

* **Personal characteristics**: `age`, `gender`, `marital_status`, `distance_from_home`
* **Job-related features**: `job_satisfaction`, `monthly_income`, `years_at_company`, `department`
* **Outcome**: `attrition` – whether the employee has left the company

This gives us a useful set of variables for exploring whether different employee groups differ in meaningful ways.

### The Method: Rule-of-Thumb and K-Means Segmentation {.unnumbered}

Our goal is to **segment employees** into groups with similar profiles and examine whether attrition differs across those groups. To do this, we will:

1. create a simple rule-of-thumb segmentation using two variables
2. compare it to K-means clustering on those same variables
3. build a richer K-means model using five scaled numeric variables
4. use the elbow method to help choose a reasonable number of clusters
5. interpret the resulting clusters using their average characteristics and attrition rates

This allows us to compare a simple hand-built segmentation to a richer data-driven one, and to ask whether the more flexible approach produces more useful insight.


### Loading the R packages and data

Here's the packages we will need to complete the exercises:

```{r}
#| message: false
library(tidyverse)  # data manipulation and plotting
library(janitor)    # cleaning column names
library(recipes)    # For data preprocessing
library(cluster)    # clustering
library(factoextra) # k-means visualization and cluster evaluation
library(broom)      # tidy output of clustering results
```


Let's load the data that we will use today:

```{r}
df <-
    read_csv("data/hr_attrition.csv") |> 
    # tidy up column names to snake_case 
    clean_names()
```

